Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-03-20 resex.tex
Revised acknowledgements. Updated corresponding author. Defined command for computer code.

\newcommand{\q}[2]{\ensuremath{#1\ \mathrm{#2}}} % quantity with units

@@

\thanks{Fermilab is operated by Fermi Research Alliance, LLC under
	Contract No.~DE-AC02-07CH11359 with the United States Department of
	Energy. This work was partially supported by the US DOE LHC
	Accelerator Research Program (LARP) and by the European FP7 HiLumi
	LHC Design Study, Grant Agreement 284404.}

\author{Miriam Fitterer}
 \email{mfittere@fnal.gov}
\author{Giulio Stancari}
\author{Alexander Valishev}
\affiliation{Fermi National Accelerator Laboratory, Batavia, Illinois, USA}

@@

\begin{table*}
  \caption{Summary of parameters used in numerical simulations of
    distribution tracking and frequency-map analysis (FMA). Further
    details can be found in Refs.~\cite{md_sim_hel_res_ex_fitterer,
      resexmd2017}.}
  \label{tab:sim_param}
  \begin{ruledtabular}
    \begin{tabular}{lcc}
      Parameter & distribution & FMA \\
      \colrule
      beam &\multicolumn{2}{c}{Beam~1} \\
      beam energy &\multicolumn{2}{c}{450 GeV} \\
      normalized emittance & \q{3.5}{\mu m} & \q{2.5}{\mu m} \\
      $4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns} \\
      $1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm} \\
      particle distribution & \parbox{0.4\textwidth}{6D Gaussian
                              distribution with $10^4$ particles}
                               & \parbox{0.4\textwidth}{equally spaced
                                 grid in $x, y$ up to $10\sigma$, with
                                 $(\Delta p / p_0) = 0$} \\
      turns tracked & $10^6$ & $10^4$ \\
      \hline
      optics & \multicolumn{2}{c}{2016 or 2017 injection optics, with
               $\beta^* = \q{11}{m}$ at IP1 and IP5} \\
      machine imperfections & standard errors, with $a_1 = b_1 =
                              0$\footnote{Orbit
                              errors are disabled due to
                              different implementation of the
                              $a_1, b_1$ coefficients in Lifetrac and
                              MAD-X. $b_2$ errors are adjusted to
                              yield an average
                              peak $\beta$-beat of 15\% over 60~seeds, as
                              observed in optics measurements
                              in the LHC.} & no errors \\
      octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MOF}} =
                   \q{+19.6}{A}$, $I_{\mathrm{MOD}} = \q{-19.6}{A}$} \\
      tunes $(Q_x, Q_y)$ & \multicolumn{2}{c}{(64.28, 59.31) for 2016,
                           (62.27, 60.295) for 2017} \\
      chromaticities $(Q'_x, Q'_y)$ & \multicolumn{2}{c}{(+15, +15)} \\
      \hline
      transverse aperture & $5.7\sigma$ &  \\
      longitudinal aperture & $10\sigma$ &  \\
    \end{tabular}
  \end{ruledtabular}
\end{table*}

For the preparation and interpretation of the experiments, two
different types of simulations were performed:
%
\begin{itemize}
\item tracking of a Gaussian distribution, referred to as
  `distribution tracking' in this paper, to obtain particle loss rates
  and emittance evolution;
\item frequency-map analysis (FMA), to visualize the location and
  intensity of resonances~\cite{fmalaskar}.
\end{itemize}
%
For both simulation types, the tracking code Lifetrac~\cite{lifetrac}
was used. The simulation parameters are summarized in
Table~\ref{tab:sim_param}. Further details are given in
Refs.~\cite{md_sim_hel_res_ex_fitterer, resexmd2017}.

@@

\begin{acknowledgments}
  We wish to acknowledge Giulia Papotti with her help with the
  preparations for the 2016 experiments and her general support from
  the operations side, Roderik Bruce and Gianluca Valentino for there
  very helpful suggestions in preparing the experiment at the LHC and
  we would like to thank all participants of the experiment for
  helping acquire the data presented in this paper. We wish to thank
  Sergei Nagaitsev for the very useful discussion and profound
  question to find out why the transverse damper is not damping any
  resonant excitation. We would like to acknowledge Dmitry Shatilov
  for his help with the Lifetrac code, and St\'{e}phane Fartoukh,
  Riccardo De Maria and Rogelio Tom\'as for their help with generating
  the appropiate optics model. We are grateful also for the support
  from the beam instrumentation team, in particular Enrico Bravin and
  Georges Trad, with the BSRT and are thankful for the collaboration
  with St\'{e}phania Papadopoulou and Fanouria Antonio for the
  analysis of the BSRT profiles.
\end{acknowledgments}

>>>>

\newcommand{\q}[2]{\ensuremath{#1\ \mathrm{#2}}} % quantity with units
\newcommand{\code}[1]{\textsc{#1}} % computer code

@@

\thanks{This manuscript has been authored by Fermi Research Alliance,
  LLC under Contract No.~DE-AC02-07CH11359 with the U.S.\ Department
  of Energy (DOE), Office of Science, Office of High Energy
  Physics. This work was partially supported by the U.S.\ DOE LHC
  Accelerator Research Program (LARP), by the European FP7 HiLumi LHC
  Design Study, Grant Agreement 284404, and by the High Luminosity LHC
  (HL-LHC) Project.}

\author{Miriam Fitterer}
\author{Giulio Stancari}
\email[Corresponding author: ]{stancari@fnal.gov}
\author{Alexander Valishev}
\affiliation{Fermi National Accelerator Laboratory, Batavia, Illinois, USA}

@@

\begin{table*}
  \caption{Summary of parameters used in numerical simulations of
    distribution tracking and frequency-map analysis (FMA). Further
    details can be found in Refs.~\cite{md_sim_hel_res_ex_fitterer,
      resexmd2017}.}
  \label{tab:sim_param}
  \begin{ruledtabular}
    \begin{tabular}{lcc}
      Parameter & distribution & FMA \\
      \colrule
      beam &\multicolumn{2}{c}{Beam~1} \\
      beam energy &\multicolumn{2}{c}{450 GeV} \\
      normalized emittance & \q{3.5}{\mu m} & \q{2.5}{\mu m} \\
      $4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns} \\
      $1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm} \\
      particle distribution & \parbox{0.4\textwidth}{6D Gaussian
                              distribution with $10^4$ particles}
                               & \parbox{0.4\textwidth}{equally spaced
                                 grid in $x, y$ up to $10\sigma$, with
                                 $(\Delta p / p_0) = 0$} \\
      turns tracked & $10^6$ & $10^4$ \\
      \hline
      optics & \multicolumn{2}{c}{2016 or 2017 injection optics, with
               $\beta^* = \q{11}{m}$ at IP1 and IP5} \\
      machine imperfections & standard errors, with $a_1 = b_1 =
                              0$\footnote{Orbit
                              errors are disabled due to
                              different implementation of the
                              $a_1, b_1$ coefficients in \code{Lifetrac} and
                              \code{MAD-X}. $b_2$ errors are adjusted to
                              yield an average
                              peak $\beta$-beat of 15\% over 60~seeds, as
                              observed in optics measurements
                              in the LHC.} & no errors \\
      octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MOF}} =
                   \q{+19.6}{A}$, $I_{\mathrm{MOD}} = \q{-19.6}{A}$} \\
      tunes $(Q_x, Q_y)$ & \multicolumn{2}{c}{(64.28, 59.31) for 2016,
                           (62.27, 60.295) for 2017} \\
      chromaticities $(Q'_x, Q'_y)$ & \multicolumn{2}{c}{(+15, +15)} \\
      \hline
      transverse aperture & $5.7\sigma$ &  \\
      longitudinal aperture & $10\sigma$ &  \\
    \end{tabular}
  \end{ruledtabular}
\end{table*}

For the preparation and interpretation of the experiments, two
different types of simulations were performed:
%
\begin{itemize}
\item tracking of a Gaussian distribution, referred to as
  `distribution tracking' in this paper, to obtain particle loss rates
  and emittance evolution;
\item frequency-map analysis (FMA), to visualize the location and
  intensity of resonances~\cite{fmalaskar}.
\end{itemize}
%
For both simulation types, the tracking code
\code{Lifetrac}~\cite{lifetrac} was used. The simulation parameters
are summarized in Table~\ref{tab:sim_param}. Further details are given
in Refs.~\cite{md_sim_hel_res_ex_fitterer, resexmd2017}.

@@

\begin{acknowledgments}
  There are several people we would like to acknowledge for their
  support, insights, participation in studies, and assistance in data
  collection. In particular, Giulia Pa\-potti offered essential
  support for machine operations and helped prepare the 2016
  experiments. Roderik Bruce and Gianluca Va\-len\-ti\-no made
  important suggestions on the design of the experiments. Dmitry
  Sha\-ti\-lov shared his deep understanding of numerical tracking
  simulations with the \code{Lifetrac} code. St\'{e}\-phane
  Far\-toukh, Ric\-car\-do De Maria and Ro\-ge\-lio Tom\'{a}s offered
  invaluable assistance in generating the appropiate accelerator
  lattice models. We wish to thank Sergei Na\-gai\-tsev for the
  insightful discussions on the interaction between the transverse
  damping system and resonant excitations. We are grateful to
  En\-ri\-co Bra\-vin, Georges Trad, and the whole beam
  instrumentation team for their crucial support on
  synchrotron-radiation diagnostics. We thank Ste\-fania
  Pa\-pa\-do\-pou\-lou and Fa\-nou\-ria An\-to\-niou for the fruitful
  collaboration on the analysis of the synchrotron-radiation profiles.

  This work was made possible by the High Luminosity LHC (HL-LHC)
  Project, the HiLumi LHC Design Study, and by the United States
  Department of Energy LHC Accelerator Research Program (LARP).
\end{acknowledgments}
